Combined annotation exploration for SCPCP000006

### 2025-02-10 analyses/cell-type-wilms-tumor-06/notebook/07_combined_annotation_across_samples_exploration.Rmd
07_annotations_v2

```{r packages, message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(patchwork)
library(DT)
```

@@

- the proportion of estimated CNV per chromosome, `{i}` in 1 to 22 `proportion_cnv_chr{i}`

@@

```{r path_to_query}
sample_metadata_file <- file.path(repository_base, "data", "current", "SCPCP000006", "single_cell_metadata.tsv")
metadata <- read.table(sample_metadata_file, sep = "\t", header = TRUE)

sample_ids <- metadata |>
  dplyr::filter(seq_unit != "spot") |>
  dplyr::pull(scpca_sample_id) |>
  unique()

@@

        # proportion of cnv per chromosome
        proportion_cnv_chr1p = srat$proportion_cnv_chr1p,
        proportion_cnv_chr2p = srat$proportion_cnv_chr2p,
        proportion_cnv_chr3p = srat$proportion_cnv_chr3p,
        proportion_cnv_chr4p = srat$proportion_cnv_chr4p,
        proportion_cnv_chr5p = srat$proportion_cnv_chr5p,
        proportion_cnv_chr6p = srat$proportion_cnv_chr6p,
        proportion_cnv_chr7p = srat$proportion_cnv_chr7p,
        proportion_cnv_chr8p = srat$proportion_cnv_chr8p,
        proportion_cnv_chr9p = srat$proportion_cnv_chr9p,
        proportion_cnv_chr10p = srat$proportion_cnv_chr10p,
        proportion_cnv_chr11p = srat$proportion_cnv_chr11p,
        proportion_cnv_chr12p = srat$proportion_cnv_chr12p,

        proportion_cnv_chr16p = srat$proportion_cnv_chr16p,
        proportion_cnv_chr17p = srat$proportion_cnv_chr17p,
        proportion_cnv_chr18p = srat$proportion_cnv_chr18p,
        proportion_cnv_chr19p = srat$proportion_cnv_chr19p,
        proportion_cnv_chr20p = srat$proportion_cnv_chr20p,

        proportion_cnv_chr1q = srat$proportion_cnv_chr1q,
        proportion_cnv_chr2q = srat$proportion_cnv_chr2q,
        proportion_cnv_chr3q = srat$proportion_cnv_chr3q,
        proportion_cnv_chr4q = srat$proportion_cnv_chr4q,
        proportion_cnv_chr5q = srat$proportion_cnv_chr5q,
        proportion_cnv_chr6q = srat$proportion_cnv_chr6q,
        proportion_cnv_chr7q = srat$proportion_cnv_chr7q,
        proportion_cnv_chr8q = srat$proportion_cnv_chr8q,
        proportion_cnv_chr9q = srat$proportion_cnv_chr9q,
        proportion_cnv_chr10q = srat$proportion_cnv_chr10q,
        proportion_cnv_chr11q = srat$proportion_cnv_chr11q,
        proportion_cnv_chr12q = srat$proportion_cnv_chr12q,
        proportion_cnv_chr13q = srat$proportion_cnv_chr13q,
        proportion_cnv_chr14q = srat$proportion_cnv_chr14q,
        proportion_cnv_chr15q = srat$proportion_cnv_chr15q,
        proportion_cnv_chr16q = srat$proportion_cnv_chr16q,
        proportion_cnv_chr17q = srat$proportion_cnv_chr17q,
        proportion_cnv_chr18q = srat$proportion_cnv_chr18q,
        proportion_cnv_chr19q = srat$proportion_cnv_chr19q,
        proportion_cnv_chr20q = srat$proportion_cnv_chr20q,
        proportion_cnv_chr21q = srat$proportion_cnv_chr21q,
        proportion_cnv_chr22q = srat$proportion_cnv_chr22q,

@@

@@

```{r fig.width=10, fig.height=4, out.width='100%'}
do_Feature_mean <- function(df, group.by, feature) {
  df <- df %>%
    group_by(sample_id, !!sym(group.by)) %>%
    summarise(m = mean(!!sym(feature)))

@@

@@

- _normal_ cells can be observe in all five compartments (`normal`,`endothelium`, `immune`, `stroma` or `fetal nephron`) and do not have CNV
We only allow a bit of flexibility in terms of CNV profile for immune and endothelium cells that have a high predicted score.
Please note that in `06_infercnv.R`, we renamed all `immune` and `endothelium` cells with a `fetal_kidney_predicted.compartment.score` > `r params$predicted.celltype.threshold` as `normal`.
Indeed, we know that false positive CNV can be observed in a cell type specific manner.

@@

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
# Define normal cells
# We first pick up the immune and endothelial cells annotated via the label transfer compartments under the condition that the predicted score is above the threshold
cell_type_df <- cell_type_df |>
  mutate(first.level_annotation = case_when(
    # assign normal/cancer based on condition
    compartment %in% c("fetal_nephron", "stroma") & 
      compartment.score > params$predicted.celltype.threshold & 
      has_cnv_score == "no CNV" ~ "normal",

@@

Using this basic strategy, we identified `r table(cell_type_df$first.level_annotation)["cancer"]` _cancer_ cells, `r table(cell_type_df$first.level_annotation)["normal"]` _normal_ cells.
Only `r table(cell_type_df$first.level_annotation)["unknown"]`cells remain _unknown_

@@

- _blastema_,: we defined as _blastema_ every cancer cell that has a `fetal_kidney_predicted.cell_type == mesenchymal cell`.
We know that these _mesenchymal_ cells are cells from the cap mesenchyme that are not expected to be in a mature kidney.
These blastema cells should express higher _CITED1_.

@@

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
cell_type_df <- cell_type_df |>
  mutate(second.level_annotation = case_when(
    # assign normal cells based on condition
    cell_type_df$compartment %in% c("fetal_nephron") &
      cell_type_df$compartment.score > params$predicted.celltype.threshold &
      cell_type_df$has_cnv_score == "no CNV" ~ "kidney",

    cell_type_df$compartment %in% c("stroma") &
      cell_type_df$compartment.score > params$predicted.celltype.threshold &
      cell_type_df$has_cnv_score == "no CNV" ~ "normal stroma",

@@

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
tmp <- cell_type_df[cell_type_df$second.level_annotation != "unknown",]
ggplot(tmp, aes(x = umap.umap_1, y = umap.umap_2, color = second.level_annotation)) +
  geom_point(size = 1, shape = 19) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))
```

#### `Second.level_annotation` of cancer cells only - `umap` reduction

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
ggplot(cell_type_df[cell_type_df$first.level_annotation == "cancer", ], aes(x = umap.umap_1, y = umap.umap_2, color = second.level_annotation)) +
  geom_point(size = 1, shape = 19) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))
```

@@

We look for each `second.level_annotation` and `sample_id`, the proportion of cells that has CNV in each of the chromosome.

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "q")))
}
```
#### gains on long arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "q")))
}
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "p")))
}
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "p")))
}
```

@@

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000081237")
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000110799")
```

#### Stroma, `Vimentin` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000026025")
```

#### Stroma, `COL6A3` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000163359")
```

#### Stroma, `THY1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000154096")
```

#### Blastema, `CITED1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000125931")
```

#### Blastema, `NCAM1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000149294")
```

#### stemness marker (blastema and primitive epithelium), `SIX2` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000170577")
```

#### Epithelium, `CDH1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000039068")
```

#### Epithelium, `PODXL` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df, group.by = "second.level_annotation", feature = "ENSG00000128567")
```

@@

- Combining label transfer and CNV inference we have produced draft annotations for all 40 Wilms tumor samples in SCPCP000006

>>>>

```{r packages, message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(patchwork)
library(DT)
library(ggpubr)
```

@@

- the CNV `{dupli|loss}` estimation per chromosome {i} and arm `{p|q}`: 
`has_{dupli|loss}_chr{i}{p|q}`, in 1 to 22

@@

We will also use:

- the samples metadata or clinical data reported in `r file.path(repository_base, "data", "current", "SCPCP000006", "single_cell_metadata.tsv")`

- the cytoband file to get the length of each chromosome arm (for plotting only)

```{r input}
# sample metadata
sample_metadata_file <- file.path(repository_base, "data", "current", "SCPCP000006", "single_cell_metadata.tsv")
metadata <- read.table(sample_metadata_file, sep = "\t", header = TRUE)

# the cytoband file that will be used to extract the size of each chromosome arm for plotting only
arm_order_file <- file.path(module_base, "results", "references", "hg38_cytoBand.txt")

```{r path_to_query}
# Extract from each Seurat object information that will be used
sample_ids <- metadata |>
  dplyr::filter(seq_unit != "spot") |>
  dplyr::pull(scpca_sample_id) |>
  unique()

@@

@@

```{r }
# Add clinical metadata
cell_type_df <- cell_type_df |>
  left_join( metadata, by= c("sample_id" = "scpca_sample_id"))

@@

```{r fig.width=10, fig.height=4, out.width='100%'}
do_Feature_mean <- function(df, group.by, feature) {
  df <- df |>
    group_by(sample_id, !!sym(group.by)) |>
    summarise(m = mean(!!sym(feature)))

@@

#### do_BoxPlot_mean

`do_BoxPlot_mean` shows boxplot of the percentage of cells with CNV in a specific chromosome arm per clinical relevant group (metadata)

- `df` is the name of the table containing metadata and feature expression (counts) per cells
- `group.by` is the name of the metadata to group the cells
- `comparison` is a list of factors in `group.by` for which we will compare the means using an unpaired Wilcoxon test

```{r fig.width=20, fig.height=15, out.width='100%', results='asis'}

do_BoxPlot_mean <- function(df, group.by, comparison){
  tmp <- df |>
    select(!!sym(group.by), sample_id, contains("has_loss_chr"), contains("has_dupli_chr")) |>
    group_by(!!sym(group.by), sample_id) |>
    summarise_all(.funs = mean) |>
    pivot_longer(cols = -c(sample_id,!!sym(group.by)) ) |>
    rename("chromosome_arm" = "name", "percentage" = "value") |>
    mutate(chromosome_arm = factor(chromosome_arm, levels = c(has_dupli_chr, has_loss_chr)))

  p <- ggplot(tmp, aes(x = !!sym(group.by), y = percentage, fill = !!sym(group.by))) +
  geom_boxplot()+
  scale_fill_brewer(palette="PuRd") +
  geom_dotplot(binaxis='y', stackdir='center') +
  facet_wrap(~chromosome_arm, nrow = 5 ) +
  theme_classic()  +
  stat_compare_means( method = "wilcox.test", paired = FALSE,  hide.ns = TRUE, label = "p.signif",  bracket.size = 0.5,tip.length = 0.01 ,comparison = list(comparison), size =8, label.y = 0.9)

  return(p)
}

@@

- _normal_ cells can be observe in all five compartments (`normal`,`endothelium`, `immune`, `stroma` or `fetal nephron`) and do not have CNV.
We only allow a bit of flexibility in terms of CNV profile for immune and endothelium cells that have a high predicted score.
Indeed, we know that false positive CNV can be observed in a cell type specific manner.
Please note that in `06_infercnv.R`, we renamed all `immune` and `endothelium` cells with a `fetal_kidney_predicted.compartment.score` > `r params$predicted.celltype.threshold` as `normal`.

@@

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
# Define normal cells
# We first pick up the immune and endothelial cells annotated via the label transfer compartments under the condition that the predicted score is above the threshold
cell_type_df <- cell_type_df |>
  mutate(first.level_annotation = case_when(
    # assign normal/cancer based on condition
    compartment %in% c("fetal_nephron", "stroma") & 
      has_cnv_score == "no CNV" ~ "normal",

@@

Using this basic strategy, we identified `r table(cell_type_df$first.level_annotation)["cancer"]` _cancer_ cells, `r table(cell_type_df$first.level_annotation)["normal"]` _normal_ cells.
`r table(cell_type_df$first.level_annotation)["unknown"]`cells remain _unknown_

@@

- _blastema_,: we defined as _blastema_ every cancer cell that has a `fetal_kidney_predicted.cell_type == mesenchymal cell`.
We know that these _mesenchymal_ cells are cells from the cap mesenchyme that are not expected to be in a mature kidney.
These blastema cells should express higher _CITED1_ and/or _NCAM1_.

@@

```{r fig.width=10, fig.height=10, out.width='100%', results='asis'}
cell_type_df <- cell_type_df |>
  mutate(second.level_annotation = case_when(
    # assign normal cells based on condition
    cell_type_df$compartment %in% c("fetal_nephron") &
      cell_type_df$has_cnv_score == "no CNV" ~ "kidney",

    cell_type_df$compartment %in% c("stroma") &
      cell_type_df$has_cnv_score == "no CNV" ~ "normal stroma",

@@

```{r fig.width=20, fig.height=20, out.width='100%', results='asis'}
cell_type_df_sub <- cell_type_df[cell_type_df$second.level_annotation != "unknown",]
ggplot(cell_type_df_sub, aes(x = umap.umap_1, y = umap.umap_2, color = second.level_annotation)) +
  geom_point(size = 1, shape = 19) +
  facet_wrap(facets = ~sample_id, ncol = 5) +
  theme_bw() +
  theme(text = element_text(size = 22))
```

@@

We look for each `second.level_annotation` and `sample_id` at the proportion of cells that has CNV in each of the chromosome.

These heatmaps allow us to:

- validate that mostly tumor cells do present high percentages of cells with CNV (by definition)

- identify chromosome and cell types that might be more sensitive to _FALSE_ positive CNV, such as the _chr6p_ which is a hub for immune genes

- get the repartition per sample of the CNV

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "q")))
}
```

#### gains on long arms

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in 1:22) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "q")))
}
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_loss_chr", i, "p")))
}
```

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
for (i in c(1:12, 16:20)) {
  print(do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = glue::glue("has_dupli_chr", i, "p")))
}
```

### Mean CNV profile

While the above heatmaps can be informative, we have to acknowledge that there are numerous and it is difficult to get a global picture of the CNV profile of the entire dataset.

Here we aim to plot a mean CNV profile across the 35 annotated samples as presented in Figure 1A of [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). This represent the mean SNP Array profile of 64 pediatric kidney cancer (including 60 Wilms tumors).

Here we plot for each chromosome arm (x-axis) the number of samples having a `dupli` (y-axis > 0, in red) or a `loss` (y-axis < 0, in blue).

For a sample, a CNV is defined as `clonal` (opaque) if detected in more than 70% of the cells;
and `subclonal` (transparent) when detected in 20 to 70% of the cells.

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}

# Here we define the vector of ordered chromosomes arms that will be used for plotting

has_dupli_chr <- c("has_dupli_chr1p",                                      
                   "has_dupli_chr1q",

                   "has_dupli_chr2p",
                   "has_dupli_chr2q",

                   "has_dupli_chr3p",
                   "has_dupli_chr3q",

                   "has_dupli_chr4p",
                   "has_dupli_chr4q",

                   "has_dupli_chr5p",
                   "has_dupli_chr5q",

                   "has_dupli_chr6p",
                   "has_dupli_chr6q",

                   "has_dupli_chr7p",
                   "has_dupli_chr7q",

                   "has_dupli_chr8p",
                   "has_dupli_chr8q",

                   "has_dupli_chr9p",
                   "has_dupli_chr9q",

                   "has_dupli_chr10p",
                   "has_dupli_chr10q",

                   "has_dupli_chr11p",
                   "has_dupli_chr11q",

                   "has_dupli_chr12p",
                   "has_dupli_chr12q",

                   "has_dupli_chr13p",
                   "has_dupli_chr13q",

                   "has_dupli_chr14p",
                   "has_dupli_chr14q",

                   "has_dupli_chr15p",
                   "has_dupli_chr15q",

                   "has_dupli_chr16p",
                   "has_dupli_chr16q",

                   "has_dupli_chr17p",
                   "has_dupli_chr17q",

                   "has_dupli_chr18p",
                   "has_dupli_chr18q",

                   "has_dupli_chr19p",
                   "has_dupli_chr19q",

                   "has_dupli_chr20p",
                   "has_dupli_chr20q",

                   "has_dupli_chr21p",
                   "has_dupli_chr21q",

                   "has_dupli_chr22p",
                   "has_dupli_chr22q"
                   )

#################################################################################

# define the vector of ordered chromosomes arms##################################
has_loss_chr <-  c("has_loss_chr1p",                                        
                   "has_loss_chr1q",

                   "has_loss_chr2p",
                   "has_loss_chr2q",

                   "has_loss_chr3p",
                   "has_loss_chr3q",

                   "has_loss_chr4p",
                   "has_loss_chr4q",

                   "has_loss_chr5p",
                   "has_loss_chr5q",

                   "has_loss_chr6p",
                   "has_loss_chr6q",

                   "has_loss_chr7p",
                   "has_loss_chr7q",

                   "has_loss_chr8p",
                   "has_loss_chr8q",

                   "has_loss_chr9p",
                   "has_loss_chr9q",

                   "has_loss_chr10p",
                   "has_loss_chr10q",

                   "has_loss_chr11p",
                   "has_loss_chr11q",

                   "has_loss_chr12p",
                   "has_loss_chr12q",

                   "has_loss_chr13p",
                   "has_loss_chr13q",

                   "has_loss_chr14p",
                   "has_loss_chr14q",

                   "has_loss_chr15p",
                   "has_loss_chr15q",

                   "has_loss_chr16p",
                   "has_loss_chr16q",

                   "has_loss_chr17p",
                   "has_loss_chr17q",

                   "has_loss_chr18p",
                   "has_loss_chr18q",

                   "has_loss_chr19p",
                   "has_loss_chr19q",

                   "has_loss_chr20p",
                   "has_loss_chr20q",

                   "has_loss_chr21p",
                   "has_loss_chr21q",

                   "has_loss_chr22p",
                   "has_loss_chr22q"
                   )

#################################################################################
# Here we define the size of each chromosome arm that will be used for plotting of the mean CNV profile

  # Load cytoBand file
  cytoBand <- read.table(gzfile(arm_order_file), header = FALSE, sep = "\t", stringsAsFactors = FALSE)
  colnames(cytoBand) <- c("chrom", "start", "end", "band", "stain")

  # Extract chromosome arm information
  cytoBand <- cytoBand |>
    mutate(arm = substr(band, 1, 1))  # Extract 'p' or 'q' from the band column

  # Define arm boundaries
  chromosome_arms <- cytoBand |>
    group_by(chrom, arm) |>
    summarize(
      Start = min(start),
      End = max(end),
      .groups = "drop"
    ) |>
    mutate(Size = End - Start,
           dupli = paste0("has_dupli_", chrom, arm),
           loss = paste0("has_loss_", chrom, arm)) |> 
  filter(arm != "")

  chromosome_arms <- chromosome_arms |>
    mutate(Size = Size / sum(Size))

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}

# prepare the table for the number of sample with a (sub)clonal loss

CNV_profile_loss <- cell_type_df_sub |>
  select(sample_id, contains("has_loss_chr")) |>
  group_by(sample_id) |>
  summarise_all(.funs = mean) |>
  pivot_longer(cols = -sample_id) |>
  rename("chromosome_arm" = "name", "percentage" = "value") |>
  left_join( chromosome_arms, by = c("chromosome_arm"="loss")) |>
  mutate(chromosome_arm = factor(chromosome_arm, levels = has_loss_chr)) |>
  mutate(absolute_70 = case_when((percentage > 0.7) ~ -1,
         .default = 0),
         absolute_10 = case_when((percentage > 0.1) ~ -1,
         .default = 0))

# prepare the table for the number of sample with a (sub)clonal gain

CNV_profile_gain <- cell_type_df_sub |>
  select(sample_id, contains("has_dupli_chr")) |>
  group_by(sample_id) |>
  summarise_all(.funs = mean) |>
  pivot_longer(cols = -sample_id) |>
  rename("chromosome_arm" = "name", "percentage" = "value") |>
  left_join( chromosome_arms, by = c("chromosome_arm"="dupli")) |>
  mutate(chromosome_arm = factor(chromosome_arm, levels = has_dupli_chr)) |>
  mutate(absolute_70 = case_when((percentage > 0.7) ~ 1,
         .default = 0),
         absolute_10 = case_when((percentage > 0.1) ~ 1,
         .default = 0))

# plot

loss_profile <- ggplot(CNV_profile_loss, aes(x = chromosome_arm, y = absolute_10, width = 20 * Size)) +
  geom_bar(stat = "identity",  fill = "#41b6c4", alpha = 0.4) +
  geom_bar(aes(x = chromosome_arm, y = absolute_70, width = 20 * Size), stat = "identity",  fill = "#41b6c4") +
  scale_x_discrete(labels = gsub("has_loss_", "", levels(droplevels(CNV_profile_loss$chromosome_arm))))  +
  theme_classic() +
  ylab("Number of occurrence") +
  xlab("chromosom arms") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 12), text = element_text(size = 15))

gain_profile <- ggplot(CNV_profile_gain, aes(chromosome_arm, y = absolute_10, width = 20* Size)) +
  geom_bar(stat = "identity",  fill = "#ce1256", alpha = 0.4) +
  geom_bar(aes(x = chromosome_arm, y = absolute_70, width = 20 * Size), stat = "identity",  fill = "#ce1256") +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        text = element_text(size = 15)) +
  ylab(" ")

# combined plot
wrap_plots(gain_profile, loss_profile, ncol = 1)

Comparing with the profile in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1), we identified important CNV known to play a role in Wilms tumor, including:

- _chr8q_, _chr12_, _chr13_, _chr18q_ gains

- _chr11q_, _chr16_, _chr17q_ loss

However, we spotted some differences:

- we couldn't detect any _chr4q_ loss as described in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).

- we identified 2 losses in _chr5_ and _chr6_ that do not show up in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
These can be `infercnv` induced _FALSE_positive CNV.

Our major concern is the pattern of the _chr1_. 
We do see a tendency of _chr1p_ loss and _chr1q_ gain but it is not as pronounced as in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
_chr1q_ gain is however one of the most promising predictive marker of Wilms tumor outcome ([Nelson et al, Curr Opin Pediatr (2020)](https://pmc.ncbi.nlm.nih.gov/articles/PMC7919941/)).
It is however important to notice that the profile shown in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1) derived from SIOP samples that have been pre-treated with chemotherapy, while samples in the dataset are enriched in `upfront resection` samples that haven't been pre-treated, according to the COG protocol. 
So far, we do not know how the pre-operative chemotherapy impact the clonal selection of specific CNV.

### Mean of the percentage of cells with CNV in a specific chromosome arm per clinical relevant group

#### `vital_status`

As described in [Nelson et al, Curr Opin Pediatr (2020)](https://pmc.ncbi.nlm.nih.gov/articles/PMC7919941/), 
[Phelps et al, Children (2018)](https://pmc.ncbi.nlm.nih.gov/articles/PMC6262554/) and 
[Zheng et al, Front. Oncol (2023)](https://www.frontiersin.org/journals/oncology/articles/10.3389/fonc.2023.1137346/full), 
LOH at _chr1p_, _chr11p_ and _chr16q_ as well as _chr1q_ gain are prognostic factors.

Here we aim to see if we identify CNV associated with poorer prognosis.

```{r fig.height=15, fig.width=20, message=FALSE, warning=FALSE, out.width='100%', results='asis'}

do_BoxPlot_mean(df = cell_type_df_sub, group.by = "vital_status", comparison = c("Alive", "Expired") )

In our study, we validate the association of _chr1q_ gain, _chr11q_ loss and _chr14q_ loss with poorer prognosis.

##### `treatment`

As we don't know the effect of the treatment on the selection of cells with CNV, it is difficult to compare our CNV profile with the one in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).
We those check the repartition of CNV in samples that have been pre-treated (i.e `Resection post chemotherapy`) with samples that have been removed surgically before any treatment (i.e `Upfront resection`).

```{r fig.height=15, fig.width=20, message=FALSE, warning=FALSE, out.width='100%', results='asis'}

do_BoxPlot_mean(df = cell_type_df_sub, group.by = "treatment", comparison = c("Upfront resection", "Resection post chemotherapy") )

Our analysis suggest that there is indeed an enrichment in cells with CNV after chemotherapy, 
including a tendency of more _chr1q_ gain samples after chemotherapy. 
This could explain some of the differences observed between our mean CNV profile and the one published in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1).

@@

Finally, we aim to validate the `second level annotation` using marker genes.

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000081237")
```

Immune cells are well annotated based on _PTPRC_ expression.

@@

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000110799")
```

Endothelial cells are well annotated based on _VWF_ expression.

#### Stroma

(*) To the best of our knowledges, there is no Wilms tumor specific marker of stroma cells.
They should express in some extend similar markers as normal stromal cells including _VIM_, _THY1_ and _COL6A3_.

##### `Vimentin` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000026025")
```

_VIM_ do not seem to be neither specific nor universal.

##### `COL6A3` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000163359")
```

_COL6A3_ expression is higher in Wilms tumor stroma cells and expressed across a wider range of samples.

##### `THY1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000154096")
```

_THY1_ do not seem to be neither specific nor universal.

#### Blastema

(*) To the best of our knowledges, there is no Wilms tumor specific and universel (i.e. expressed by all cells in every sample) marker of blastema cells.
We expect however blastema cells to express higher levels of stemness markers such as _CITED1_, _SIX1/2_, _NCAM1_.

##### `CITED1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000125931")
```

Blastema cells are enriched in _CITED1_+ cells, but some samples do not express it at all.

##### `NCAM1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000149294")
```

_NCAM1_ seems to be more cancer specific and broadly expressed among samples.

##### `SIX2` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000170577")
```

#### Epithelium

##### `CDH1` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000039068")
```

As expected, epithelial cells express higher expression of _CDH1_.

##### `PODXL` expression

```{r fig.width=20, fig.height=5, out.width='100%', results='asis'}
do_Feature_mean(cell_type_df_sub, group.by = "second.level_annotation", feature = "ENSG00000128567")
```

@@

- Combining label transfer and CNV inference we have produced draft annotations for 35/40 Wilms tumor samples in SCPCP000006.
We decided not to annotate samples for which the `infercnv` couldn't be run with a normal reference, as we don't trust the results in that case and want to avoid misclassification (`SCPCS000177`, `SCPCS000180`, `SCPCS000181`, `SCPCS000190`, `SCPCS000197`).

- The mean CNV profile of the Wilms tumor cohort seems reasonable in regards to the one published in [Cresswell et al, BioRXiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
We identify some commonalities but also some divergence. 
They can be due to technical limitation of our CNV inference but can also reflect some specificities of this cohort (e.g. pre/post treatment samples).
It would be really useful to compare for some sample the `infercnv` result to a ground truth such as SNP Array (to be requested).
